%\documentclass{sig-alternate-sigmod08}
\documentclass[10pt,conference]{IEEEtran} 
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{moreverb}
\usepackage{cancel}
\usepackage{fancyhdr}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{epstopdf}
\usepackage{amssymb,amsfonts,amsmath,amsthm}
%\usepackage{tikz}
%\usepackage{tikz,fullpage,tikz_custom}
\usepackage{subfigure}
%\usepackage{multicol}
%\usepackage{coordsys}
%\usetikzlibrary{arrows,%
%	          	petri,%
%					topaths}%
%\usepackage{tkz-berge}
%\usepackage[position=top]{subfig}
%\usepackage{epstopdf}
%\usepackage{amsmath}


%%%%%%%%%%  NEEDS to be commente out to have the nice headers %%%%%
%\pagestyle{fancy}
\DeclareGraphicsRule{.tic}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

%\textwidth = 6.5 in
%\textheight = 9 in
%\oddsidemargin = 0.0 in
%\evensidemargin = 0.0 in
%\topmargin = 0.0 in
%\headheight = 0.0 in
%\headsep = 0.0 in
%\parskip = 0.2in
%\parindent = 0.0in
%\abovedisplayskip
%\belowdisplayskip
 
%Rnd{\classname}{FOO 699}
\newcommand{\doc}{Technical Report}
\newcommand{\doctitle}{Efficient Recovery from False State in Distributed Routing Algorithms}
\newcommand{\myname}{Daniel Gyllstrom}
%\newcommand{\qed}{\hfill $\Box$ \hfill \\}


\newtheorem{theorem}{Theorem}
\setcounter{theorem}{0}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{meta-proposition}[theorem]{Meta-Proposition}
\newtheorem{note}[theorem]{Note}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{proviso}[theorem]{Proviso}         
\newtheorem{question}[theorem]{Question}         
\newtheorem{remark}[theorem]{Remark}         
\newtheorem{define}{Definition}     

\newcommand{\minv}{$\overrightarrow{min}$ }
\newcommand{\minvi}{$\overrightarrow{min}_i$ }
\newcommand{\minvis}{$\overrightarrow{min}_i$}
\newcommand{\minvj}{$\overrightarrow{min}_j$ }
\newcommand{\minvjs}{$\overrightarrow{min}_j$}
\newcommand{\minvv}{$\overrightarrow{min}_v$ }
\newcommand{\minvvs}{$\overrightarrow{min}_v$}
\newcommand{\dmatrix}{$dmatrix$ }
\newcommand{\dmatrixs}{$dmatrix$}
\newcommand{\dmatrixi}{$dmatrix_i$ }
\newcommand{\dmatrixis}{$dmatrix_i$} 
\newcommand{\dmatrixj}{$dmatrix_j$ }
\newcommand{\dmatrixjs}{$dmatrix_j$} 
\newcommand{\dmatrixv}{$dmatrix_v$ }
\newcommand{\dmatrixvs}{$dmatrix_v$} 
%\newcommand{\dmatrixvs}{{\tt dmatrix$_v$}}
\newcommand{\dv}{{DV$^+$ }}
%\newcommand{\bad}{{$\hat{v}$ }}
\newcommand{\Bad}{{$\overline{V}$ }}
\newcommand{\Bads}{{$\overline{V}$}}
\newcommand{\bad}{{$\overline{v}$ }}
\newcommand{\bads}{{$\overline{v}$}}
\newcommand{\alg}{{DV$^+$ }}
\newcommand{\block}{{\tt todo-rename }}
\newcommand{\second}{{{\tt 2}$^{{\tt nd}}$ {\tt best} }}
\newcommand{\seconds}{{{\tt 2}$^{{\tt nd}}$ {\tt best}}}
\newcommand{\infinity}{{count-to-$\infty$ }}
\newcommand{\purge}{{{\tt purge} }}
\newcommand{\purges}{{{\tt purge}}}
\newcommand{\resetall}{{{\tt reset-all} }}
\newcommand{\resetalls}{{{\tt reset-all}}}
\newcommand{\resetk}{{{\tt reset-k} }}
\newcommand{\resetks}{{{\tt reset-k}}}
%\newcommand{\badvector}{{$v_{bad-lie}$ }}
%\newcommand{\oldvector}{{$v_{bad-old}$ }}
\newcommand{\badvector}{$\overrightarrow{bad}$ }
\newcommand{\badvectors}{$\overrightarrow{bad}$}
\newcommand{\oldvector}{$\overrightarrow{old}$ }
\newcommand{\oldvectors}{$\overrightarrow{old}$}
\newcommand{\finalvector}{{$v_{final}$ }}
\newcommand{\illigit}{{illegitimate path} }
\newcommand{\illigits}{{illegitimate paths} }
\newcommand{\lcd}{$\Delta_{lc}$ }
\newcommand{\lcds}{$\Delta_{lc}$s }
\newcommand{\cpr}{{\tt cpr} }
\newcommand{\cprs}{{\tt cpr}}
\newcommand{\er}{Erd\"{o}s-R\'enyi }
\newcommand{\ers}{Erd\"{o}s-R\'enyi} 
\newcommand{\reword}{{\it  Comment: reword}. }
\newcommand{\more}{{\it  Comment: more details needed here}. }
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

%newcommand{\block}{{\textsc{block-wait} }}


%%%%%%%%%%%%% This adds the header to each page - currently a bit broken so commented out
%\lhead{\myname{}}
%\chead{\myname{}}
%\rhead{\doc{}}
%\lfoot{}
%

\makeatletter
\newcommand{\un}[1]{%
   % \ifmmode \@@underline{#1} \else %
             $\@@underline{\hbox{#1}}$\fi}
\makeatother
\raggedbottom

\begin{document}




\title{\doctitle}


\author{
\IEEEauthorblockN{Daniel Gyllstrom, Sudarshan Vasudevan, Jim Kurose, Gerome Miklau } 
\IEEEauthorblockA{Department of Computer Science \\ 
University of Massachusetts Amherst \\ 
\{dpg, svasu, kurose, miklau\}@cs.umass.edu}} 
%\andI 
%\IEEEauthorblockN{Sudarshan Vasudevan} 
%\IEEEauthorblockA{Department of Computer Science \\ 
%University of Massachusetts Amherst \\ 
%svasu@cs.umass.edu} 
%\and
%\IEEEauthorblockN{James Kurose} 
%\IEEEauthorblockA{Department of Computer Science \\ 
%University of Massachusetts Amherst \\ 
%kurose@cs.umass.edu} }

\maketitle

\input{abstract}

\section{Introduction}
\label{sec:intro}

%Distributed systems are vulnerable to malicious and misconfigured nodes which inject false system state. 

Malicious and misconfigured nodes can degrade the performance of a distributed system by injecting incorrect state information. Such false state can then be further propagated 
through the system either directly in its original form or indirectly, e.g., as a result of diffusing computations initially using this false state.  In this paper, we consider 
the problem of removing such false state from a distributed system.

In order to make the false-state-removal problem concrete, we investigate distance vector routing as an instance of this problem. Distance vector forms the basis for many routing 
algorithms widely used in the Internet (e.g., BGP, a path-vector algorithm) and in multi-hop wireless networks (e.g., AODV, diffusion routing). However, distance vector is vulnerable 
to compromised nodes that can potentially flood a network with false routing information, resulting in erroneous least cost paths, packet loss, and congestion. Such scenarios have occurred
in practice. For example, in 1997 a significant portion of Internet traffic was routed through a single misconfigured router, rendering a large part of the Internet inoperable for several
hours \cite{Neumann97}. More recently \cite{Google}, a routing error forced Google to redirect its traffic through Asia, causing congestion that left many Google services unreachable. 
Distance vector currently has no mechanism to recover from such scenarios. Instead, human operators are left to manually reconfigure routers. It is in this context that we propose and
evaluate automated solutions for recovery.

In this paper, we design, develop, and evaluate three different approaches for correctly recovering from the injection of false routing state (e.g., a compromised node incorrectly
claiming a distance of $0$ to all destinations). Such false state, in turn, may propagate to other routers through the normal execution of distance vector routing, making this
a network-wide problem. Recovery is correct if the routing tables in all nodes have converged to a global state in which all nodes have removed each compromised node as a destination,
and no node bears a least cost path to any destination that routes through a compromised node.
%none of that node’s least cost paths to any destination route through a compromised node. 

Specifically, we develop three novel distributed recovery algorithms: \seconds, \purges, and \cprs. \second performs localized state invalidation, followed by network-wide recovery. 
Nodes directly adjacent to a compromised node locally select alternate paths that avoid the compromised node; the traditional distributed distance vector algorithm is then executed to 
remove remaining false state using these new distance vectors. The \purge algorithm performs global false state invalidation by using diffusing computations to invalidate distance vector 
entries (network-wide) that routed through a compromised node. As in \seconds, traditional distance vector routing is then used to recompute distance vectors. \cpr uses local snapshots 
and a rollback mechanism to implement recovery. Although our solutions are tailored to distance vector routing, we believe they represent approaches that are applicable to other instances
of this problem.

We prove the correctness of each algorithm, derive communication complexity bounds for each algorithm, and use simulations to evaluate the efficiency of each algorithm in
terms of message overhead and convergence time.
Our simulations show that \cpr using poison reverse outperforms \second and \purge (with and without poison reverse) -- at the cost of checkpoint memory --
over topologies with fixed and changing link costs. This is because \cpr efficiently removes all false state by rolling back to a checkpoint
immediately preceding the injection of false routing state. 
In scenarios where link costs can change, \purge using poison reverse yields performance close to \cpr with poison reverse. \purge makes 
use of computations subsequent to the injection of false routing state that do not depend on false routing state, while \cpr 
must process all valid link cost changes that occurred since false routing state was injected.  Finally, our simulations show that poison reverse significantly improves 
performance for all three algorithms, especially for topologies with changing link costs. % by removing pairwise routing loops.

Recovery from false routing state is closely related to the problem of
recovering from malicious transactions \cite{Liu98,Liu00} in
distributed databases. Our problem is also similar to that of rollback
in optimistic parallel simulation \cite{Jeff}. However, we are unaware
of any existing solutions to the problem of recovering from false
routing state. A closely related problem to the one considered in this
paper is that of discovering misconfigured nodes. In
Section~\ref{sec:problem}, we discuss existing solutions to this
problem. In fact, the output of these algorithms serve as input to the
recovery algorithms proposed in this paper.

This paper has eight sections. In Section \ref{sec:problem} we define the problem and state our assumptions.
We present our three recovery algorithms in Section \ref{sec:algs} and prove the correctness of each algorithm in Section \ref{sec:correct}.
Then, in Section \ref{sec:analysis}, we present a qualitative evaluation 
of our recovery algorithms and derive communication complexity bounds for each algorithm. 
Section \ref{sec:eval} describes our simulation study. We detail related work in Section \ref{sec:related} and finally we conclude and 
comment on directions for future work in Section \ref{sec:future}.







\input{problem-stmt}


\input{algorithms}


\input{correct}

\input{analysis}

\input{experiments}

\input{related}

\input{future}

\section{Acknowledgments}
The authors greatly appreciate discussions with Dr. Brian DeCleene of BAE Systems, who initially suggested this problem area.



\bibliographystyle{plain}
\bibliography{tech-report2}

\input{appendix}


\end{document}
