\section{Problem Formulation}
\label{sec:problem}

We consider distance vector routing \cite{Gall87} over arbitrary network topologies. We model a network as an undirected graph, $G=(V,E)$,
with a link weight function $w: E \rightarrow \mathbb{N}$.
Each node, $v$, maintains the following state as part of distance vector: a vector of all adjacent nodes ($adj(v)$), a vector of least cost distances to all
nodes in $G$ (\minvvs), and a \emph{distance matrix} that contains distances to every node in the network via each adjacent node (\dmatrixvs). 

We assume that the identity of the compromised node is provided by a different algorithm, and thus do not consider this problem in this paper.
Examples of such algorithms include \cite{Arini, Feam, Feldmann}
in the context of wired networks and \cite{Paul02} in the wireless setting. 
Specifically, we assume that at time $t$, this algorithm is used to notify all neighbors of the 
compromised node(s). Let $t'$ be the time the node was compromised.

For each of our algorithms, the goal is for all nodes to recover ``correctly'': all nodes should remove the compromised node as a destination and find
new least cost distances that do not use the compromised node. If the network becomes disconnected as a result of removing the compromised node, all
nodes need only compute new least cost distances to all other nodes within their connected component.

For simplicity, let \bad denote the compromised node, let \oldvector refer to $\overrightarrow{min}_{\overline{v}}$ 
before \bad was compromised, and let \badvector denote $\overrightarrow{min}_{\overline{v}}$ after \bad has been compromised.


